Likelihood-free estimation of contact network parameters from viral phylogenies

### 2016-06-29 doc/defence/transtree.tex
working on defence talk

\documentclass{beamer}
\usepackage{tikz}
\usepackage{gillius}
\usepackage{abraces}
\usepackage{etex}
\usepackage{array}
\usepackage{multirow}
\usepackage[backend=biber, style=numeric-comp, sorting=none]{biblatex}

% http://tex.stackexchange.com/questions/12703/how-to-create-fixed-width-table-columns-with-text-raggedright-centered-raggedlef
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

\usetikzlibrary{external}
\usetikzlibrary{shapes}
\usetikzlibrary{arrows}
\usetikzlibrary{positioning}
\usetikzlibrary{decorations.pathreplacing}
\usetikzlibrary{calc}
\usetikzlibrary{automata}

\usetheme[everytitleformat=regular]{m}
\setbeamertemplate{navigation symbols}{}

\graphicspath{{../figures/}}
\newcommand{\tablepath}{../tables}

\title{Phylogenetic inference of contact network parameters with approximate Bayesian computation}
\author[RMM \& AFYP]{Rosemary M McCloskey$^1$ \and Bioinformatics Training Program \\ Supervisor: Art FY Poon$^{1,2}$}
\institute[UBC \& BCCfE]{$^1$BC Centre for Excellence in HIV/AIDS, Vancouver, Canada \\ $^2$Department of Medicine, University of British Columbia, Vancouver, Canada}
\date{M.Sc. Defence, July 26, 2016}

\newcommand{\dd}[2]{\frac{\text{d}\,#1}{\text{d}\,#2}}

\addbibresource{papers.bib}

\begin{document}
\setbeamercolor{background canvas}{bg=white}
\renewcommand{\footnotesize}{\tiny}
\definecolor{red}{RGB}{228,26,28}
\definecolor{blue}{RGB}{55,126,184}
\definecolor{green}{RGB}{77,175,74}
\definecolor{purple}{RGB}{152,78,163}

\begin{frame}{Transmission trees shape viral phylogenies}
  \begin{center}
    \begin{tikzpicture}[
        every node/.style = {inner sep=0pt},
        every path/.style = {->, >=stealth, very thick},
        label/.style = {anchor=south, color=white, inner sep=4pt}
      ]
      \def\pht{1cm}
      \node (abcd) {\includegraphics[height=\pht]{stock/person}};
      \node at (abcd.south) [label] {\Large a};

@@

      \coordinate [below right=0.5 and 3 of b] (abcd);
      \coordinate [below right=1 and 1 of abcd] (b);
      \coordinate [above left=1 and 0.25 of abcd] (acd);
      \coordinate [above left=1 and 0.25 of acd] (a);
      \coordinate [right=of acd] (cd);
      \coordinate [right=1.5 of cd] (c);
      \coordinate [above right=of cd] (d);

      \uncover<1-2>{
      \node (x) at (abcd) {\includegraphics[height=\pht]{stock/virus}};
      \node at (x.center) {\large a};
      }

      \uncover<2>{
      \node (x) [below right=0.3 of abcd, anchor=north west] {\includegraphics[height=\pht]{stock/virus}}; 
      \node at (x.center) {\large b};
      }

      \uncover<3-4>{
      \node (x) [above=0 of acd, anchor=south] {\includegraphics[height=\pht]{stock/virus}}; 
      \node at (x.center) {\large a};
      }

      \uncover<3->{
      \node (x) [below right=-0.2 of b, anchor=north west] {\includegraphics[height=\pht]{stock/virus}}; 
      \node at (x.center) {\large b};

      \draw (abcd) -- (acd);
      \draw (abcd) -- (b);
      }

      \uncover<4>{
      \node (x) [right=0.3 of acd, anchor=west] {\includegraphics[height=\pht]{stock/virus}}; 
      \node at (x.center) {\large c};
      }

      \uncover<5-6>{
      \node (x) at (cd) [anchor=west] {\includegraphics[height=\pht]{stock/virus}};  
      \node at (x.center) {\large c};
      }

      \uncover<5->{
      \node (x) at (a) [anchor=south] {\includegraphics[height=\pht]{stock/virus}};
      \node at (x.center) {\large a};

      \draw (acd) -- (cd);
      \draw (acd) -- (a);
      }

      \uncover<6>{
      \node (x) [above right=0.4\pht and 0.4\pht of cd] [anchor=south west] {\includegraphics[height=\pht]{stock/virus}};
      \node at (x.center) {\large d};
      }

      \uncover<7->{
      \node (x) at (c) [anchor=west] {\includegraphics[height=\pht]{stock/virus}};
      \node at (x.center) {\large c};

      \node (x) at (d) [anchor=south west] {\includegraphics[height=\pht]{stock/virus}};
      \node at (x.center) {\large d};

      \draw (cd) -- (c);
      \draw (cd) -- (d);
      }

    \end{tikzpicture}
  \end{center}
\end{frame}

\end{document}

>>>>

\begin{frame}{Contact networks shape transmission trees}
  \vspace{-1cm}
  \begin{center}
    \begin{tikzpicture}
      [every node/.style = {inner sep=0pt},
       every path/.style = {->, >=stealth, very thick},
       label/.style = {anchor=south, color=white, inner sep=4pt},
       time/.style = {anchor=north west, inner sep=2pt}]
      \def\pht{1cm}
      \node (a) {\includegraphics[height=\pht]{stock/person}};
      \node at (a.south) [label] {\Large a};
      \node [below left=0.75 of a] (b) {\includegraphics[height=\pht]{stock/person}};
      \node at (b.south) [label] {\Large b};
      \node [below=0.75 of b] (e) {\includegraphics[height=\pht]{stock/person}};
      \node at (e.south) [label] {\Large e};
      \node [below right=0.75 of a] (c) {\includegraphics[height=\pht]{stock/person}};
      \node at (c.south) [label] {\Large c};
      \node [below=0.75 of c] (d) {\includegraphics[height=\pht]{stock/person}};
      \node at (d.south) [label] {\Large d};

      \fill [shade, opacity=0.2] (a.east) -- (c.north) -- (c.north west) -- (a.south east) -- cycle;
      \fill [shade, opacity=0.2] (a.west) -- (b.north) -- (b.north east) -- (a.south west) -- cycle;
      \fill [shade, opacity=0.2] (a.south) -- (a.south west) -- (e.north) -- (e.north east) -- cycle;
      \fill [shade, opacity=0.2] (b.south west) -- (b.south) -- (e.north) -- (e.north west) -- cycle;
      \fill [shade, opacity=0.2] ($(c.south) + (-8pt, 0)$) -- ++(16pt, 0) -- ($(d.north) + (8pt, 0)$) -- ++(-16pt, 0) -- cycle;
      \fill [shade, opacity=0.2] (b.east) -- (b.south east) -- (d.west) -- (d.north west) -- cycle;
      \uncover<2->{
      \draw [red] ($(a.west) + (0, -12pt)$) -- ($(b.north) + (8pt, 0)$);
      }
      \uncover<4->{
      \draw [blue] ($(a.east) + (0, -12pt)$) -- ($(c.north) + (-8pt, 0)$);
      }
      \uncover<6->{
      \draw [green!80!black] (c) -- (d);
      }

      \node (abcd) [above right=0 and 5cm of a] {\includegraphics[height=\pht]{stock/person}};
      \node at (abcd.south) [label] {\Large a};

@@

      \coordinate [above right=0 and 2.5cm of abcd] (top);
      \draw (top) -- ++ (0, -6cm) node [anchor=west, inner xsep=4pt] {time};

    \end{tikzpicture}

  $\qquad$contact network \hfill transmission tree$\qquad$
  \end{center}
\end{frame}

### 2016-06-29 doc/defence/virusphylo.tex
working on defence talk

\documentclass{beamer}
\usepackage{tikz}
\usepackage{gillius}
\usepackage{abraces}
\usepackage{etex}
\usepackage{array}
\usepackage{multirow}
\usepackage[backend=biber, style=numeric-comp, sorting=none]{biblatex}

% http://tex.stackexchange.com/questions/12703/how-to-create-fixed-width-table-columns-with-text-raggedright-centered-raggedlef
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

\usetikzlibrary{external}
\usetikzlibrary{shapes}
\usetikzlibrary{arrows}
\usetikzlibrary{positioning}
\usetikzlibrary{decorations.pathreplacing}
\usetikzlibrary{calc}
\usetikzlibrary{automata}

\usetheme[everytitleformat=regular]{m}
\setbeamertemplate{navigation symbols}{}

\graphicspath{{../figures/}}
\newcommand{\tablepath}{../tables}

\title{Phylogenetic inference of contact network parameters with approximate Bayesian computation}
\author[RMM \& AFYP]{Rosemary M McCloskey$^1$ \and Bioinformatics Training Program \\ Supervisor: Art FY Poon$^{1,2}$}
\institute[UBC \& BCCfE]{$^1$BC Centre for Excellence in HIV/AIDS, Vancouver, Canada \\ $^2$Department of Medicine, University of British Columbia, Vancouver, Canada}
\date{M.Sc. Defence, July 26, 2016}

\newcommand{\dd}[2]{\frac{\text{d}\,#1}{\text{d}\,#2}}

\addbibresource{papers.bib}

\begin{document}
\setbeamercolor{background canvas}{bg=white}
\renewcommand{\footnotesize}{\tiny}
\definecolor{red}{RGB}{228,26,28}
\definecolor{blue}{RGB}{55,126,184}
\definecolor{green}{RGB}{77,175,74}
\definecolor{purple}{RGB}{152,78,163}

@@

      \draw (cd) -- (c);
      \draw (cd) -- (d);
      }

    \end{tikzpicture}
  \end{center}
\end{frame}

\end{document}

>>>>

@@

      \draw (cd) -- (c);
      \draw (cd) -- (d);
      }
    \end{tikzpicture}
  \end{center}
\end{frame}

### 2016-06-29 doc/ismb/poster.tex
working on defence talk

\title{Phylogenetic estimation of contact network parameters with kernel-ABC}
\author{Rosemary M. McCloskey$^1$ and Art F. Y. Poon $^{1,2}$}
\date{March 11, 2015}
\institute{$^1$ BC Centre for Excellence in HIV/AIDS 
           $^2$ Department of Medicine, University of British Columbia}
\begin{document}

\node at (current page.west) 
  [anchor=north] {
  \includegraphics[width=\paperwidth, trim=0 0 0 1in]{logos/cfe_banner}
};

@@

% hacky things
\node [text width=6in] at ($ (current page.west) + (-0.42\paperwidth, -1.5in) $)
  { \LARGE B.I.G. Research Day \\ March 11, 2016 \\ Vancouver, Canada};
\node [text width=6in, align=right] at ($ (current page.west) + (0.42\paperwidth, -1.5in) $)
  { \LARGE rmccloskey@cfenet.ubc.ca \\ (604) 682-6344 ext. 63266 };

@@

  \block{Barab\'asi-Albert preferential attachment networks}{
    Nodes are added one-at-a-time until there are $N$ total. Each time a node
    is added, $m$ new edges are connected to it. Each new edge is attached to
    an existing node of degree $d$ with probability proportional to
    $d^\alpha$.

    \vspace{1cm}
    \centerline{\includegraphics[width=0.2\textwidth, trim=0 22in 0 0, clip]{alpha-bounds}}

@@

    \vspace{0.5cm}
    \captionof{table}{
    Three networks/trees were simulated under each parameter combination.
    Kernel-ABC was used to re-estimate the parameters.
    }
  }
  \block{Accurate estimates for $\alpha$, $I$; not $m$, $N$}{
    \centerline{\includegraphics[width=0.16\textwidth]{abc-point-estimate}}
    Dotted lines are true values. $\alpha$ and $I$, which had the most impact
    on tree shape, were possible to estimate accurately. Posterior
    distributions of $m$ and $N$ heavily resemble the priors, indicating
    little information about these parameters in the trees.
  }
  \block{Narrow dispersion of $\alpha$, $I$ compared to $m$, $N$}{
    \centerline{\includegraphics[width=0.16\textwidth]{abc-posterior-example}}
  }

@@

  \block{Transmission trees were estimated from viral sequences}{
    \centerline{\includegraphics[width=0.2\textwidth]{pipeline}}

@@

  \block{Parameter heterogeneity among real datasets}{
    \centerline{\includegraphics[width=0.2\textwidth]{realdata-hpd}}
  }
\end{columns}

>>>>

\title{Likelihood-free estimation of contact network parameters from viral phylogenies}
\author{Rosemary M. McCloskey$^1$, Richard H. Liang$^1$, and Art F. Y. Poon $^{1,2}$}
\date{March 11, 2015}
\institute{$^1$ BC Centre for Excellence in HIV/AIDS 
           $^2$ Department of Medicine, University of British Columbia}
\begin{document}

\colorlet{blockbodybgcolor}{white}
\node at (current page.west) 
  [anchor=north] {
  \includegraphics[width=\paperwidth, trim=0 0 0 1in]{logos/cfe_banner}
};

@@

% hacky things
\node [text width=6in] at ($ (current page.west) + (-0.42\paperwidth, -1.5in) $)
  { \LARGE ISMB \\ July 8-12, 2016 \\ Orlando, USA};
\node [text width=6in, align=right] at ($ (current page.west) + (0.42\paperwidth, -1.5in) $)
  { \LARGE rmccloskey@cfenet.ubc.ca \\ (604) 682-6344 ext. 63266 };

@@

  \block{Barab\'asi-Albert preferential attachment networks}{
    Nodes are added one at a time until there are $N$ total. Each time a node
    is added, $m$ new edges are connected to it. Each new edge is attached to
    an existing node of degree $d$ with probability proportional to
    $d^\alpha$.

    \vspace{1cm}
    \centerline{\includegraphics[width=0.2\textwidth]{pa-example}}
    \vspace{1cm}

@@

    \vspace{0.5cm}
    \captionof{table}{
    Three networks/trees were simulated under each parameter combination.
    Kernel-ABC was used to re-estimate the parameters.
    }
  }
  \block{Accurate estimates for $\alpha$, $I$; not $m$, $N$}{
    \centerline{\includegraphics[width=0.16\textwidth]{abc-point-estimate-m2}}
    Dotted lines are true values. $\alpha$ and $I$, which had the most impact
    on tree shape, were possible to estimate accurately. Posterior
    distributions of $m$ and $N$ heavily resemble the priors, indicating
    little information about these parameters in the trees.
  }
  \block{Narrow dispersion of $\alpha$, $I$ compared to $m$, $N$}{
    \centerline{\includegraphics[width=0.16\textwidth]{abc-posterior-example}}
  }

@@

  \block{Transmission trees were estimated from viral sequences}{
    \vspace{1cm}
    \centerline{\includegraphics[width=0.25\textwidth]{pipeline}}
    \vspace{1cm}

@@

  \block{Parameter heterogeneity among real datasets}{
    \vspace{1cm}
    \centerline{\includegraphics[width=0.25\textwidth]{realdata-hpd}}
    \vspace{1cm}
  }
\end{columns}
